### 2020-08-25 LaTeX/sections/paper.tex
Paper - until LL vs LLMapping

In this research work, we design and implement a call market on the Ethereum blockchain. In our design, time is treated as discrete and not continuous and orders are accumulated over predetermined time intervals and are processed in batch instead of serially in order of receipt. The trading day is divided into frequent but discrete time intervals; every time the market opens traders submit their ask and bid order messages. These orders are accumulated in a sorted data structure and when the trading period is over, orders are matched against each other.  Our proposed decentralized exchange is both data structure and token agnostic (traders can trade any token that is ERC20 compliant). The upcoming orders are not stored in the Call Market contract, instead they are stored in a priority queue where each element is associated with a priority and is served according to its priority. A priority queue can be implemented using many of the data structures that differ in efficiency and costs. Our main \textit{Call Market} smart contract is written in 268 lines of Solidity, a high level programming language that is syntactically similar to Java~\cite{Ethereum41:online}, using the Truffle development framework and deployed on Ganache-CLI. We used Javascript for testing by leveraging the Mocha testing framework. To facilitate a safe exchange among buyers and sellers, we implement the Call Market smart contract in the form of a \textit{collateralized}; for market participants to be able to send bid and/or ask orders, they have to first supply tokens (depending on what ERC20 tokens they aim to trade) as collaterals by calling any of the  \texttt{DepositToken()} or \texttt{DepositEther()} methods. The collateralized Call Market acts as a payment guarantee so traders cannot default on payment or delivery of their assets.

@@

As mentioned in~\ref{sec:callmarket}, a priority queue can be implemented using different data structures. We implemented five different generic priority queues that support four main functions (see Table~\ref{tab:PQ_API}). Followings describe each implementation.

% = = = = = = = = = = = =Operations for a generic Priority Queue Table = = = = = = = = = = = = =  %
\begin{table}[]
\centering
\begin{tabular}{|c|c|}
\hline

@@

% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{Enqueue()}       				& inserts an element into the priority queue                        \\ \hline
	\textbf{Dequeue()}					& removes and returns the highest priority element 		\\ \hline
	\textbf{isEmpty()}						& checks if the priority queue is empty 					\\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %

\end{tabular}
\caption{\footnotesize{Operations for a generic Priority Queue}
\label{tab:PQ_API}}
\end{table}
% = = = = = = = = = = = = = = = = = = = = =  = = = = = = = = = = =  = = = = = = = = = == = = =  %

% = = = = = = = = = = = = = = = = = = = = =  = = =% 
\subsection{Heap with Dynamic Array}

A Heap is a type of binary tree data structure that comes in two forms of a (i) Max-Heap and (ii) Min-Heap. All the nodes of tree are in a specific order and the root of the tree always represents the highest priority item of the data structure (the largest and smallest values in the Max-Heap and Min-Heap respectively). We implemented a priority queue with a Heap that itself is implemented using Solidity dynamically-sized array.  
% = = = = = = = = = = = = = = = = = = = = =  = = =%

\subsection{Heap with Static Array}

%Here, we implemented a priority queue using the Heap that itself is stored in Solidity static arrays. Every time the priority queue contract i

We enter 50 unsigned integers to the priority queues in random ordering. To do so, we use JavaScript \texttt{Math.random()} function to generate pseudo-random integers between 1 and 200. Figure~\ref{fig:average_uints_insertion} shows the gas cost variations for entering 50 unsigned integers in the five data structures. The x-axis is the place in line (\eg the10th number entered in the priority queue) and the and y-axis is the cost of that transaction in gas.

% = = = = = = = = = = = =average_uints_insertion Figure = = = = = = = = = = = = =  %

\begin{figure}[htb!p]
\centering
\includegraphics[width=1\textwidth]{fig/average_uints_insertion.png}
\caption{\footnotesize{}  \label{fig:average_uints_insertion}}
\end{figure}

@@

Then, we call the \texttt{Dequeue()} function which iteratively removes the highest priority number from the priority queue (until the data structure is empty). The computational costs for dequeuing 50 unsigned integers in each priority queue are outlined in Table~\ref{tab:PQ_UnitTests}. The tests are performed using the current Ethereum gas metrics (block gas limit $=$11,741,495 and 1 gas $=$ 56 gwei)~\footnote{https://ethstats.net/}. The second column of the table shows the net gas consumption (the gasUsed value derived from transaction receipts) for removing 50 integers from each priority queues.

At the time of this writing, Ethereum transaction receipts only contain the net gas consumption and not the total gas consumption ( total gas consumption is defined as $gas refunded + gasUsed$ ) and we cannot find out the value of the EVM's refund counter from inside the EVM.

So in order to account for refunds inside each priority queue smart contract, we can calculate them manually; first we figure out exactly how much storage is being cleared when dequeuing  the max integers and then we could multiply the number of storage slots cleared by 15,000 (see the last column of Table~\ref{tab:PQ_UnitTests}).

Another way to know the amount of refund in each priority queue is to use the \texttt{estimateGas API} which provides a rough idea about the total amount of gas that is required for a transaction to go through. The \texttt{web3.eth.estimateGas} pretends the transaction is included in the block and its functions (with the parameters passed) will be executed on the Ethereum blockchain. Doing so, it provides us an estimate of how much gas is needed to be sent with the transaction. The second and third columns of Table~\ref{tab:PQ_UnitTests}) show the total amount of gas required for dequeuing 50 integers from each priority queue (provided by estimateGas) and the amount of gas refund ($Total Gas Consumption - gasUsed$) respectively.

Note that in order to urge miners to process smart contract with refunds, the accumulated gas refund can never exceed half the gas used up during computation~\cite{wood2014ethereum}. So at the end of a successful transaction, the amount of gas in the refund counter (capped at half the net gas used) is returned to the caller. For example, the amount of gas that has been used when dequeuing 50 integers from the linkedlist with mapping data structure is 731,514 and since $3,000,000 > 731,514/2$, the amount of refund returned to the caller is $731,514/2 = 365,757$.

\section{Unit Testing the Priority Queues}

Here we execute the same JavaScrip test on the five priority queues with an end goal of unit testing them. We enter 50 unsigned integers to the priority queues in random ordering. To do so, we use JavaScript \texttt{Math.random()} function to generate pseudo-random integers between 1 and 200. Figure~\ref{fig:average_uints_insertion} shows the gas cost variations for entering 50 unsigned integers in the five data structures. The x-axis is the place in line (\eg the10th number entered in the priority queue) and the and y-axis is the cost of that transaction in gas.

% = = = = = = = = = = = =average_uints_insertion Figure = = = = = = = = = = = = =  %

\begin{figure}[htb!p]
\centering
\includegraphics[width=1\textwidth]{fig/average_uints_insertion.png}
\caption{\footnotesize{}  \label{fig:average_uints_insertion}}
\end{figure}

Then, we call the \texttt{Dequeue()} function which iteratively removes the maximum value of the priority queue (until the data structure is empty). The computational costs for dequeuing 50 unsigned integers in each priority queue are outlined in Table~\ref{tab:PQ_UnitTests}. The tests are performed using the current Ethereum gas metrics (block gas limit $=$11,741,495 and 1 gas $=$ 56 gwei)~\footnote{https://ethstats.net/}. The second column of the table shows the net gas consumption (the gasUsed value derived from transaction receipts) for removing 50 integers from each priority queues.

At the time of this writing, Ethereum transaction receipts only contain the net gas consumption and not the total gas consumption ( total gas consumption is defined as $gas refunded + gasUsed$ ) and we cannot find out the value of the EVM's refund counter from inside the EVM.

So in order to account for refunds inside each priority queue smart contract, we can calculate them manually; first we figure out exactly how much storage is being cleared when dequeuing  the max integers and then we could multiply the number of storage slots cleared by 15,000 (see the last column of Table~\ref{tab:PQ_UnitTests}).

Another way to know the amount of refund in each priority queue is to use the \texttt{estimateGas API} which provides a rough idea about the total amount of gas that is required for a transaction to go through. The \texttt{web3.eth.estimateGas} pretends the transaction is included in the block and its functions (with the parameters passed) will be executed on the Ethereum blockchain. Doing so, it provides us an estimate of how much gas is needed to be sent with the transaction. The second and third columns of Table~\ref{tab:PQ_UnitTests}) show the total amount of gas required for dequeuing 50 integers from each priority queue (provided by estimateGas) and the amount of gas refund ($Total Gas Consumption - gasUsed$) respectively.

Note that in order to urge miners to process smart contract with refunds, the accumulated gas refund can never exceed half the gas used up during computation~\cite{wood2014ethereum}. So at the end of a successful transaction, the amount of gas in the refund counter (capped at half the net gas used) is returned to the caller. For example, the amount of gas that has been used when dequeuing 50 integers from the linkedlist with mapping data structure is 731,514 and since $3,000,000 > 731,514/2$, the amount of refund returned to the caller is $731,514/2 = 365,757$.

% = = = = = = = = = = = =Uint testing the PQs Table with 50 Random Integers between 1 and 200 = = = = = = = = = = = = =  %

@@

\textbf{\thead{Priority Queue}}    & \textbf{\thead{Net Cost\\in Gas}}      & \textbf{\thead{Total Cost\\in Gas\\(from estimateGas)}}      & \textbf{\thead{Gas Refund \\(from \texttt{estimateGas})}}    & \textbf{\thead{Gas Refund \\(Manually Calculated)}} \\ \hline

% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{\thead{Heap with \\ Dynamic Array}}         				& 2,575,997               & 3,349,746		& 773,749             & 750,000                       \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{\thead{Heap with \\ Static Array}}           				& 1,324,856                & 2,090,182     	& 765,326             & 750,000                      \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{\thead{Mapping with \\ keys stored \\ in Heap}} 		& 2,863,239                & 4,378,584       	& 1,515,345           & 1,500,000                     \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{Linkedlist}                       							& 557,085               	& 1,772,085      	& 1,215,000           & 1,200,000                      \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{\thead{Linkedlist with \\ Mapping}}          				& 731,514              	& 3,731,514       	& 3,000,000     	  &  3,765,000                       \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %

\end{tabular}
\caption{\footnotesize{PQUnitTests with 50 Random Integers between 1 and 200}
\label{tab:PQUnitTests_random}}
\end{table}
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  %

% = = = = = = = = = = = =Uint testing the PQs Table with 50 Integers (NOT RANDOM!) = = = = = = = = = = = = =  %

\textbf{\thead{Priority Queue}}    & \textbf{\thead{Net Cost\\in Gas}}      & \textbf{\thead{Total Cost\\in Gas\\(from estimateGas)}}      & \textbf{\thead{Gas Refund \\(from \texttt{estimateGas})}}    & \textbf{\thead{Gas Refund \\(Manually Calculated)}} \\ \hline

% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{\thead{Heap with \\ Dynamic Array}}         				& 2,394,202                & 3,159,551           &765,349	    & 750,000                     \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{\thead{Heap with \\ Static Array}}           				& 1,317,922                & 2,083,240   	& 765,318          & 750,000                     \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{\thead{Mapping with \\ keys stored \\ in Heap}} 		& 2,629,191                & 4,144,536       	&1,515,345          & 1,500,000                     \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{Linkedlist}                       							&557,085            		& 1,772,085      	&1,215,000          & 1,200,000                      \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{\thead{Linkedlist with \\ Mapping}}          				&731,514              	&3,731,514      	& 3,000,000    	  &  3,765,000                       \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %

\end{tabular}
\caption{\footnotesize{PQUnitTests}
\label{tab:PQ_UnitTests}}
\end{table}
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  %

 In order to choose the best priority queue ()

Note that we have to carefully choose a proper implementation as they have different performance and costs. In order to examine 
In order to examine the cost and performanmc

%In order to enable our Call Market contract to act as an interface with any priority queue, we implemented generic types of priority queues that support four main functions (see Table~\ref{tab:PQ_API}). Followings describe each implementation.

@@

% = = = = = = = = = = = =Uint testing the PQs Table = = = = = = = = = = = = =  %

\section{Unit Testing the Priority Queues}

Here we execute the same JavaScrip test on the five priority queues with an end goal of unit testing them. We enter 50 unsigned integers to the priority queues in random ordering. To do so, we use JavaScript \texttt{Math.random()} function to generate pseudo-random integers between 1 and 200. Figure~\ref{fig:average_uints_insertion} shows the gas cost variations for entering 50 unsigned integers in the five data structures. The x-axis is the place in line (\eg the10th number entered in the priority queue) and the and y-axis is the cost of that transaction in gas.

% = = = = = = = = = = = =average_uints_insertion Figure = = = = = = = = = = = = =  %

\begin{figure}[htb!p]
\centering
\includegraphics[width=1\textwidth]{fig/average_uints_insertion.png}
\caption{\footnotesize{}  \label{fig:average_uints_insertion}}
\end{figure}

Then, we call the \texttt{Dequeue()} function which iteratively removes the maximum value of the priority queue (until the data structure is empty). The computational costs for dequeuing 50 unsigned integers in each priority queue are outlined in Table~\ref{tab:PQ_UnitTests}. The tests are performed using the current Ethereum gas metrics (block gas limit $=$11,741,495 and 1 gas $=$ 56 gwei)~\footnote{https://ethstats.net/}. The second column of the table shows the net gas consumption (the gasUsed value derived from transaction receipts) for removing 50 integers from each priority queues.

At the time of this writing, Ethereum transaction receipts only contain the net gas consumption and not the total gas consumption ( total gas consumption is defined as $gas refunded + gasUsed$ ) and we cannot find out the value of the EVM's refund counter from inside the EVM.

So in order to account for refunds inside each priority queue smart contract, we can calculate them manually; first we figure out exactly how much storage is being cleared when dequeuing  the max integers and then we could multiply the number of storage slots cleared by 15,000 (see the last column of Table~\ref{tab:PQ_UnitTests}).

Another way to know the amount of refund in each priority queue is to use the \texttt{estimateGas API} which provides a rough idea about the total amount of gas that is required for a transaction to go through. The \texttt{web3.eth.estimateGas} pretends the transaction is included in the block and its functions (with the parameters passed) will be executed on the Ethereum blockchain. Doing so, it provides us an estimate of how much gas is needed to be sent with the transaction. The second and third columns of Table~\ref{tab:PQ_UnitTests}) show the total amount of gas required for dequeuing 50 integers from each priority queue (provided by estimateGas) and the amount of gas refund ($Total Gas Consumption - gasUsed$) respectively.

Note that in order to urge miners to process smart contract with refunds, the accumulated gas refund can never exceed half the gas used up during computation~\cite{wood2014ethereum}. So at the end of a successful transaction, the amount of gas in the refund counter (capped at half the net gas used) is returned to the caller. For example, the amount of gas that has been used when dequeuing 50 integers from the linkedlist with mapping data structure is 731,514 and since $3,000,000 > 731,514/2$, the amount of refund returned to the caller is $731,514/2 = 365,757$.

% = = = = = = = = = = = =Uint testing the PQs Table with 50 Random Integers between 1 and 200 = = = = = = = = = = = = =  %

\textbf{\thead{Priority Queue}}    & \textbf{\thead{Net Cost\\in Gas}}      & \textbf{\thead{Total Cost\\in Gas\\(from estimateGas)}}      & \textbf{\thead{Gas Refund \\(from \texttt{estimateGas})}}    & \textbf{\thead{Gas Refund \\(Manually Calculated)}} \\ \hline

% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{\thead{Heap with \\ Dynamic Array}}         				& 2,575,997               & 3,349,746		& 773,749             & 750,000                       \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{\thead{Heap with \\ Static Array}}           				& 1,324,856                & 2,090,182     	& 765,326             & 750,000                      \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{\thead{Mapping with \\ keys stored \\ in Heap}} 		& 2,863,239                & 4,378,584       	& 1,515,345           & 1,500,000                     \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{Linkedlist}                       							& 557,085               	& 1,772,085      	& 1,215,000           & 1,200,000                      \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{\thead{Linkedlist with \\ Mapping}}          				& 731,514              	& 3,731,514       	& 3,000,000     	  &  3,765,000                       \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %

\end{tabular}
\caption{\footnotesize{PQUnitTests with 50 Random Integers between 1 and 200}
\label{tab:PQUnitTests_random}}
\end{table}
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  %

% = = = = = = = = = = = =Uint testing the PQs Table with 50 Integers (NOT RANDOM!) = = = = = = = = = = = = =  %

\textbf{\thead{Priority Queue}}    & \textbf{\thead{Net Cost\\in Gas}}      & \textbf{\thead{Total Cost\\in Gas\\(from estimateGas)}}      & \textbf{\thead{Gas Refund \\(from \texttt{estimateGas})}}    & \textbf{\thead{Gas Refund \\(Manually Calculated)}} \\ \hline

% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{\thead{Heap with \\ Dynamic Array}}         				& 2,394,202                & 3,159,551           &765,349	    & 750,000                     \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{\thead{Heap with \\ Static Array}}           				& 1,317,922                & 2,083,240   	& 765,318          & 750,000                     \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{\thead{Mapping with \\ keys stored \\ in Heap}} 		& 2,629,191                & 4,144,536       	&1,515,345          & 1,500,000                     \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{Linkedlist}                       							&557,085            		& 1,772,085      	&1,215,000          & 1,200,000                      \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{\thead{Linkedlist with \\ Mapping}}          				&731,514              	&3,731,514      	& 3,000,000    	  &  3,765,000                       \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %

\end{tabular}
\caption{\footnotesize{PQUnitTests}
\label{tab:PQ_UnitTests}}
\end{table}
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  %

>>>>

In this research work, we design and implement a call market on the Ethereum blockchain. In our design, time is treated as discrete and not continuous and orders are accumulated over predetermined time intervals and are processed in batch instead of serially in order of receipt. The trading day is divided into frequent but discrete time intervals; every time the market opens traders submit their ask and bid order messages. These orders are accumulated in a sorted data structure and when the trading period is over, orders are matched against each other.  Our proposed decentralized exchange is both data structure and token agnostic (traders can trade any token that is ERC20 compliant). The upcoming orders are not stored in the Call Market contract, instead they are stored in a priority queue where each element is associated with a priority and is served according to its priority. Our main \textit{Call Market} smart contract is written in 268 lines of Solidity, a high level programming language that is syntactically similar to Java~\cite{Ethereum41:online}, using the Truffle development framework and deployed on Ganache-CLI. To facilitate a safe exchange among buyers and sellers, we implement the Call Market smart contract in the form of a \textit{collateralized}; for market participants to be able to send bid and/or ask orders, they have to first supply tokens (depending on what ERC20 tokens they aim to trade) as collaterals by calling any of the  \texttt{DepositToken()} or \texttt{DepositEther()} methods. The collateralized Call Market acts as a payment guarantee so traders cannot default on payment or delivery of their assets.

@@

 A priority queue can be implemented using many of the data structures that differ in performance and costs. In order to choose the appropriate data structure for better performance and costs, we implemented and examined five different generic priority queues that support four main functions (see Table~\ref{tab:PQ_API}). Followings describe each implementation.

% = = = = = = = = = = = =Operations for a generic Priority Queue Table = = =  = = = = = = = = = = = %
\begin{table}[]
\centering
\begin{tabular}{|c|c|}
\hline

@@

% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{Enqueue()}       	& inserts an element into the priority queue                        \\ \hline
	\textbf{Dequeue()}		& removes and returns the highest priority element 		\\ \hline
	\textbf{isEmpty()}			& checks if the priority queue is empty 					\\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  %

\end{tabular}
\caption{\footnotesize{Operations for a generic Priority Queue}
\label{tab:PQ_API}}
\end{table}
% = = = = = = = = = = = = = = = = = = = = =  = = = = = = = = = = =  = = = = = = = = = == = = =  = =%

\item\textbf{{Heap with Dynamic Array.}} A Heap is a type of binary tree data structure that comes in two forms of a (i) Max-Heap and (ii) Min-Heap. All the nodes of tree are in a specific order and the root of the tree always represents the highest priority item of the data structure (the largest and smallest values in the Max-Heap and Min-Heap respectively). We implemented a priority queue with a Heap that itself is represented by Solidity dynamically-sized array.

\item\textbf{{Heap with Static Array.}} A Heap can be also represented by a Solidity storage array in which the storage is statically allocated. To do this, we pass the required sized of the array as a smart contract constructor parameter.

\item\textbf{{Mapping with keys stored in Heap.}} This priority queue stores items inside a Solidity mapping that maps the mapping keys (256-bit integers) to these items. The mapping keys are then stored (based on the value they are mapped to) in a sorted Heap that is implemented with a dynamically-sized array.

\item\textbf{{Linked List.}} Here we used a sorted singly Linked List to implement the priority queue. A Linked List is a linear data structure made of a chain of items (also known as nodes) where each item contains a value and a pointer to the next node in the list. In our implementation, each node is represented by a smart contract that contains the data and a pointer to its next smart contract in the list.

\item\textbf{{Linked List with Mapping.}} Here we used Solidity mapping to implement our priority queue with a sorted doubly Linked List where every node points to it previous \textbf{and} next node in the list. The Linked List entries are first wrapped in an struct data type that contains (i) the unique id of that element, (ii) the unique id of its previous item in the list, and (iii) the unique id of its next item in the list. These structs are then stored in the mapping indexed by item id.

Next, we used Javascript for testing by leveraging the Mocha testing framework and performed our priority queue performance tests. We entered 50 pseudo-random integers between 1 and 200 to the priority queues and obtained the associated gas costs. Figure~\ref{fig:random_insertion} represents the gas costs variations for entering 50 unsigned integers to the five priority queues; the x-axis is the place in line (\eg the10th number entered in the priority queue) and the and y-axis is the cost of that transaction in gas.

% = = = = = = = = = = = =random_insertion Figure = = = = = = = = = = = = =  %

\begin{figure}[htb!p]
\centering
\includegraphics[width=1\textwidth]{fig/random_insertion.png}
\caption{\footnotesize{Gas costs for entering 50 items into five different priority queues.}  \label{fig:random_insertion}}
\end{figure}

@@

Then, we called the \texttt{Dequeue()} function which iteratively removes the highest priority number from the priority queues (until the data structure is empty). The computational costs for dequeuing 50 unsigned integers in each priority queue are outlined in Table~\ref{tab:PQUnitTests}. These tests are performed using the current Ethereum gas metrics (block gas limit $=$11,741,495 and 1 gas $=$ 56 gwei)~\footnote{https://ethstats.net/}. The second column of the table shows the net gas consumption (the gasUsed value derived from transaction receipts) for removing 50 integers from each priority queues.

% = = = = = = = = = = = =Gas costs and refunds for Dequeueing 50 integers from each PQ  = = = = = = = = = = = = =  %

@@

\textbf{\thead{Priority Queue}}    & \textbf{\thead{Net Cost\\in Gas}}      & \textbf{\thead{Total Cost\\in Gas\\(From \texttt{estimateGas})}}      & \textbf{The Difference}    & \textbf{\thead{Gas Refund \\(Manually Calculated)}} \\ \hline

% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{\thead{Heap with \\ Dynamic Array}}         				& 2,518,131               & 3,283,480		& 765,349             & 750,000                       \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{\thead{Heap with \\ Static Array}}           				& 1,385,307                & 2,150,625     	& 765,318             & 750,000                      \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{\thead{Mapping with \\ keys stored \\ in Heap}} 		& 2,781,684                & 4,316,229       	& 1,534,545           & 1,500,000                     \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{Linked List}                       							& 557,085               	& 1,772,085      	& 1,215,000           & 1,200,000                      \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{\thead{Linked List with \\ Mapping}}          				& 731,514              	& 3,731,514       	& 3,000,000     	  &  3,765,000                       \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %

\end{tabular}
\caption{\footnotesize{The gas metrics associated with dequeuing 50 items from five priority queues.}
\label{tab:PQUnitTests}}
\end{table}
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  %

At the time of this writing, Ethereum transaction receipts only contain the net gas consumption and not the total gas consumption (total gas consumption is defined as $gas refunded + gasUsed$) and we cannot obtain the value of the EVM's refund counter from inside the EVM. So in order to account for refunds inside each priority queue smart contract, we can calculate them manually; first we figure out exactly how much storage is being cleared when dequeuing the highest priority elements and then we multiply the number of storage slots cleared by 15,000 (see the last column of Table~\ref{tab:PQUnitTests}).

Another way to obtain the amount of refund in each priority queue is to use the \texttt{estimateGas API} which provides a rough idea about the total amount of gas that is required for a transaction to go through. The \texttt{web3.eth.estimateGas} pretends the transaction is included in the block and its functions (with the parameters passed) will be executed on the Ethereum blockchain. Doing so, it provides us an estimate of how much gas is needed to be sent with the transaction. The third and fourth columns of Table~\ref{tab:PQUnitTests}) show the total amount of gas required for dequeuing 50 integers from each priority queue (provided by estimateGas) and the amount of gas refund ($Total Gas Consumption - gasUsed$) respectively.

Note that in order to urge miners to process smart contract with refunds, the accumulated gas refund can never exceed half the gas used up during computation~\cite{wood2014ethereum}. So at the end of a successful transaction, the amount of gas in the refund counter (capped at half the net gas used) is returned to the caller. For example, the amount of gas that has been used when dequeuing 50 integers from the linked list with mapping priority queue is 731,514 and since $3,000,000 > 731,514/2$, the amount of refund returned to the caller is $731,514/2 = 365,757$.

Based on our tests results, the \textit{linked list} and \textit{linked list with mapping} priority queues show the best performance in compared to the different heap implementations. In the next section, we explore the differences between the two different implementations of the linked list priority queue.

% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  %

\subsection{Linked List \textit{vs.} Linked List with Mapping}

% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  %
\begin{table}[]
\begin{tabular}{|c|c|c|c|c|}
\hline
\multirow{2}{*}{Gas Metrics}                        							& \multicolumn{2}{c|}{Linked List}           						& \multicolumn{2}{c|}{Linked List with Mapping}     					\\ \cline{2-5} 
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  = = = = = =  = = = = = = = = = = = = = = = =  = = = = = =  = = = = = = %
                                                                						& with \texttt{Self-destruct} 		& without \texttt{Self-destruct}                	& with \texttt{delete} 		& without \texttt{delete}              		\\ \hline                                                                                                                                                                                                                                                          % = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  = = = = = =  = = = = = = = = = = = = = = = =  = = = = = =  = = = = = = %
                   Net Cost in Gas                          					& 557,085                    			& 721,370                     				& 731,514                     	& 334,689                      				\\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  = = = = = =  = = = = = = = = = = = = = = = =  = = = = = =  = = = = = = %
\thead{Total Cost in Gas \\ (from \texttt{estimateGas})} 		& 1,772,085                  			& 736,370                  					& 3,731,514                     	& 1,099,689                       				\\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  = = = = = =  = = = = = = = = = = = = = = = =  = = = = = =  = = = = = = %
		    The Difference                                             		& 1,215,000                  			& 15,000                      				& 3,000,000                    	& 765,000                       				\\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  = = = = = =  = = = = = = = = = = = = = = = =  = = = = = =  = = = = = = %

\end{tabular}
\end{table}
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  %

@@
